---
title: "Meade 4000 Filter"
author: "Ralph Rogge (RRO)"
date: "17. Februar 2016"
output:
    html_document:
        fig_width: 10
        fig_height: 6
---

Set this flag to TRUE when you want to write out plots.
```{r}
write.plot.enabled = F
```

```{r, message=FALSE}
library(data.table)
library(dplyr)
library(ggplot2)
library(jsonlite)
```

```{r approx.y, echo=F}
approx.y <- function(x, y = NULL, xout, rule=2, ...) {
  approx(x,y=y,xout=xout,rule=rule,...)$y
}
```

```{r, filter.names, echo=F}
filter.names <- c(
  "8"  , "11" , "21", "23A", "25A", "38A", "47" , "56" , "58", "80A", "82A"
)
```

```{r, filter.color, echo=F}
filter.colors <- c(
  "yellow", "yellowgreen", "orange", "red", "red3", "blue", "purple", "green",  "green3", "deepskyblue3",
  "skyblue",  "black"
)
```

```{r filter.labels, echo=F}
filter.labels <- c(
  "#8", "#11", "#21", "#23A", "#25A", "#38A", "#47", "#56", "#58", "#80A", "#82A", "None"
)
```

```{r make.jupiter.data, echo=F}
make.jupiter.data <- function(filter, vision, vision.type, neb, ntrz) {
  
  subvision <- subset(vision, type==vision.type)
  wavelength <- seq(min(filter$wavelength), max(filter$wavelength))
  
  # Create data frame with wavelength, transmission, filter.
  df <- data.frame(wavelength = filter$wavelength, transmission = filter$transmission, filter = filter$name)
  
  # Add plain vision data.
  df <- rbind(df, data.frame(wavelength = wavelength, transmission = 100, filter = "None"))
  
  # Add relative luminous efficiency.
  df$rle <- approx.y(subvision$wavelength, subvision$rle, df$wavelength)

  # Add filtered luminous efficiency.
  df$fle <- df$rle * df$transmission / 100

  # Add Jupiter NEB reflectance.
  df$neb <- approx.y(neb$wavelength, neb$reflectance, df$wavelength)

  # Add Jupiter NTrZ reflectance.
  df$ntrz <- approx.y(ntrz$wavelength, ntrz$reflectance, df$wavelength)

  # Add vision.
  df$vision <- vision.type

  df
}
```

```{r make.sunlit.data, echo=F}
make.sunlit.data <- function(filter, vision, vision.type, sun, reflectance=1) {
  
  subvision <- subset(vision, type==vision.type)
  wavelength <- seq(min(filter$wavelength), max(filter$wavelength))
  
  # Create data frame with wavelength, transmission, filter.
  df <- data.frame(wavelength = filter$wavelength, transmission = filter$transmission, filter = filter$name)

  # Add plain vision data.
  df <- rbind(df, data.frame(wavelength = wavelength, transmission = 100, filter = "None"))

  # Add relative luminous efficiency.
  df$rle <- approx.y(subvision$wavelength, subvision$rle, df$wavelength)

  # Add filtered luminous efficiency
  df$fle <- df$rle * df$transmission/100 

  # Add solar spectral irrandiance
  df$irradiance <- approx.y(sun$wavelength, sun$irradiance, df$wavelength)

  # Add reflectance for a spectral uniform reflector.
  df$reflectance <- reflectance

  # Add vision.
  df$vision <- vision.type

  # Return.
  df
}
```

```{r preprocess.filter, echo=F}
preprocess.filter <- function(df, wavelength, name) {
  data.frame(
    wavelength = wavelength, 
    transmission = approx.y(df$wavelength, df$transmission, wavelength), 
    name=name
  )
}
```

```{r read.filter, echo=F}
read.filter <- function(name) {
  read.csv(name, col.names=c("wavelength","transmission","X"))[,1:2]
}
```

```{r read.vision, echo=F}
read.vision <- function(filename, type) {
  df <- read.csv(filename)
  df$type <- type
  df
}
```

```{r write.plot, echo=F}
write.plot <- function(p, name=NULL, width=1024, height=576, font.size=16) {
  title <- ifelse(!is.null(name), name, p$labels$title)
  if (is.null(title)) stop("Neither name nor title given")
  filename <- paste0("Meade 4000 - ", title, ".png")
  png(filename,width, height)
  print(p + theme_bw(base_size=font.size))
  dev.off()
}
```

## Filter Transmission

Read Meade 4000 filter set data from files
```{r}
filter <- data.frame(wavelength=c(), transmission=c(), name=c())
for (name in filter.names) {
  filename <- paste0("Meade 4000 Filter ",name,".csv")
  filter <- rbind(filter, preprocess.filter(read.filter(filename), seq(200,900), name))
}
```

Plot transmission of Meade 4000 filter set
```{r}
p <- ggplot(data=filter) +
  geom_line(aes(wavelength,transmission,color=factor(name, levels=c(filter.names, "None")))) +
  scale_colour_manual(name="Filter", values=filter.colors, labels=filter.labels) +
  labs(title="Meade 4000 Filter", x="Wavelength [nm]", y="Transmission [%]") +
  theme_bw() +
  coord_cartesian(xlim=c(250,800))
p

if (write.plot.enabled) write.plot(p, "Transmission")
```

## Scotopic/Photopic Vision

Read data for scotopic and photopic vision from file
```{r}
vision <- rbind(read.vision("scvle.csv", "Scotopic"), read.vision("vljve.csv", "Photopic"))
vision$type = factor(vision$type, levels=c("Scotopic", "Photopic")) # Reorder vision type
```

Wavelength limits for plots
```{r}
w.min <- max(min(vision$wavelength), 380)
w.max <- min(max(vision$wavelength), 720)
```

Plot relative luminous efficency of photopic and scotopic vision
```{r}
p <- ggplot(data=vision) +
  geom_line(aes(wavelength,rle,color=type)) +
  scale_color_manual(name="Vision", values=c("black","grey70")) +
  labs(title="Vision Efficiency", x="Wavelength [nm]", y="Relative Luminous Efficiency") +
  coord_cartesian(xlim=c(w.min,w.max)) +
  theme_bw()
p

if (write.plot.enabled) write.plot(p)
```

## Visual Observing of Sunlit Object

Read solar spectral distribution from file
```{r}
sun <- read.csv("ASTMG173.csv", header=F, skip=2)[,c(1,3)]
colnames(sun) <- c("wavelength","irradiance")
```

Set the average solar irradiance between 630 and 690 nm to one
```{r}
N <- mean(sun$irradiance[which(sun$wavelength == 630):which(sun$wavelength == 690)])
sun$irradiance <- sun$irradiance / N
```

Plot solar spectral distribution
```{r}
p <- ggplot(data=sun) +
  geom_line(aes(wavelength,irradiance)) +
  labs(title="Solar Spectral Irradiance / Air Mass 1.5", x="Wavelength [nm]", y="Irradiance") +
  theme_bw() +
  coord_cartesian(xlim=c(w.min,w.max))
p
```

Calculate flux for visual observing of sunlit object under scotopic and photopic conditions
```{r}
sunlit.data <- rbind(
  make.sunlit.data(filter, vision, "Scotopic", sun), 
  make.sunlit.data(filter, vision, "Photopic", sun)
)

sunlit.flux <- sunlit.data %>%

  # Reorder vision.
  mutate(vision = factor(vision, levels=c("Scotopic","Photopic"))) %>%
    
  # Group by vision and filter.
  group_by(vision, filter) %>%
  
  # Add vision specific constant.
  mutate(k = ifelse(vision == "Scotopic", 1699, 683)) %>%
  
  # Calculate flux.
  summarize(flux = sum(fle * irradiance * reflectance * k)) %>%
  
  # Sort by flux.
  arrange(-flux)

print(data.table(sunlit.flux))
```

Plot flux for visual observing of a sunlit object
```{r}
p <- ggplot(sunlit.flux) +
  geom_bar(aes(x=filter,y=flux,fill=factor(filter, levels=c(filter.names, "None"))),stat="identity") +
  scale_fill_manual(name="Filter", values=filter.colors, labels=filter.labels) +
  facet_grid(. ~ vision) +
  labs(title="Sunlit Object", x="Wavelength [nm]", y="Flux [p.d.u.]") +
  theme_bw()
p

if (write.plot.enabled) write.plot(p)
```

## Visual Observing of Jupiter

Read Jupiter NEB/NTrZ reflectance data from file
```{r}
neb <- read.csv("NEB.csv")
ntrz <- read.csv("NTrZ.csv")
```

Plot of Jupiter NEB/NTrZ reflectance
```{r}
p <- ggplot(data=rbind(neb,ntrz)) +
  geom_line(aes(wavelength, reflectance, color=feature)) +
  scale_color_discrete(name="Feature") +
  labs(title="Jupiter Reflectance", x="Wavelength [nm]", y="Reflectance [p.d.u.]") +
  theme_bw() +
  coord_cartesian(xlim=c(w.min,w.max))
p

if (write.plot.enabled) write.plot(p)
```

Calculate flux and contrast for visual observing Jupiter under scotopic and photopic conditions
```{r}
jupiter.data <- rbind(
  make.jupiter.data(filter, vision, "Scotopic", neb, ntrz), 
  make.jupiter.data(filter, vision, "Photopic", neb, ntrz)
)

jupiter.flux <- jupiter.data %>%
  
  # Adjust vision order.
  mutate(vision=factor(vision, levels=c("Scotopic","Photopic"))) %>%
  
    # Add vision specific constant.
  mutate(k = ifelse(vision=="Scotopic",1699,683)) %>%

  # Group by vision and filter.
  group_by(vision, filter) %>%
  
  # Calculate flux.
  summarize(flux.neb=sum(k*fle*neb), flux.ntrz=sum(k*fle*ntrz)) %>%
  
  # Calculate Michelson contrast.
  mutate(michelson=(flux.ntrz-flux.neb)/(flux.ntrz+flux.neb)) %>%
  
  # Calculate Weber contrast.
  mutate(weber=(flux.ntrz-flux.neb)/(flux.ntrz)) %>%
  
  # Sort by contrast.
  arrange(-michelson)

print(data.table(jupiter.flux))
```

Plot contrast for visual observing Jupiter 
```{r}
p <- ggplot(jupiter.flux) +
  geom_bar(aes(x=filter,y=michelson,fill=factor(filter, levels=c(filter.names, "None"))),stat="identity") +
  scale_fill_manual(name="Filter", values=filter.colors, labels=filter.labels) +
  facet_grid(. ~ vision) +
  labs(title="Jupiter", x="Wavelength [nm]", y="Contrast") +
  theme_bw()
p

if (write.plot.enabled) write.plot(p)
```

## Appendix

Smart approx function
```{r approx.y, echo=T}
```

Filter names
```{r filter.names, echo=T}
```

Filter colors
```{r filter.color, echo=T}
```

Filter labels
```{r filter.labels, echo=T}
```

Function to set up data frame for visual observing Jupiter
```{r make.jupiter.data, echo=T}
```

Function to set up data frame for visual observing of a sunlight object
```{r make.sunlit.data, echo=T}
```

Function to preprocess filter data
```{r preprocess.filter, echo=T}
```

Function to read in filter data
```{r read.filter, echo=T}
```

Function to read vision data from file
```{r read.vision, echo=T}
```

Function used to write plot to PNG file
```{r write.plot, echo=T}
```
